---
title: "STATS 4310 - Tuan Pham - Finance Application"
author: "Tuan Pham"
date: "2023-10-25"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Part1:**

</b>
</b>

## Part 1a:

</b>

**Stock pick:**

</b>

|Underlying Stock | Ticker  | Beta  | PE Ratio  | Growth Estimate(5yrs) |
|:----------------|:--------|:------|:----------|:----------------------|
|Microsoft        |MFST     |N/A    |32.91      |14.30%                 |
|Lockheed Martin  |LMT      |0.62   |16.41      |10.69%                 |
|Enphase Energy   |ENPH     |1.51   |23.91      |27.60%                 |


</b>

## Part 1b:

</b>

```{r message=FALSE, warning=FALSE}
# Quantitative Financial Modelling Framework
library(quantmod)
```

</b>

```{r}
# Set the time frame of desired observation 
start<- as.Date("2019-01-01")
end<-as.Date("2023-11-13")
```

</b>

```{r}
# get desired tickers by getSymbols function
MSFT <- getSymbols("MSFT",from=start, to=end, auto.assign=F)
LMT <- getSymbols("LMT",from=start, to=end, auto.assign=F)
ENPH <- getSymbols("ENPH",from=start, to=end, auto.assign=F)

```

</b>

```{r}
## Close price
MSFT.close <- MSFT[ ,4]
LMT.close <- LMT[ ,4]
ENPH.close <- ENPH[ ,4]
```

</b>

```{r}
# Daily Return function - The difference between open and close price of an underlying stock
MSFT.return <- dailyReturn(MSFT.close)
LMT.return <- dailyReturn(LMT.close)
ENPH.return <- dailyReturn(ENPH.close)

```



`Microsoft`

```{r}
plot(MSFT.close, type='l', col='green', main = "Microsoft")
```

`Lockheed Martin Corp.`

```{r}
plot(LMT.close, type='l', col='blue', main = "Lockheed Martin")
```

`Enphase Energy`

```{r}
plot(ENPH.close, type='l', col='red', main = "Enphase Energy")

```

</b>

## Part 1c:

</b>

`Annualized Expected Return and Annualized Risk`

```{r}
## Microsoft
MSFT.exp.annual.return <- mean(MSFT.return) * 252
MSFT.exp.annual.risk <- sd(MSFT.return) * sqrt(252)
MSFT.rr <- c(MSFT.exp.annual.return, MSFT.exp.annual.risk)
MSFT.rr
```

</b>

> Microsoft(MSFT) has annualized expected return of 31.42% and annualized risk of 30.79%

```{r}
## Lockheed Martin
LMT.exp.annual.return <- mean(LMT.return) * 252
LMT.exp.annual.risk <- sd(LMT.return) * sqrt(252)
LMT.rr <- c(LMT.exp.annual.return, LMT.exp.annual.risk)
LMT.rr
```

> Lockheed Martin(LMT) has annualized expected return of 14.17% and annualized risk of 26.38%

</b>

```{r}
## Enphase Energy
ENPH.exp.annual.return <- mean(ENPH.return) * 252
ENPH.exp.annual.risk <- sd(ENPH.return) * sqrt(252)
ENPH.rr <- c(ENPH.exp.annual.return, ENPH.exp.annual.risk)
ENPH.rr
```
</b>

> Enphase Energy(ENPH)  has annualized expected return of 86.36% and annualized risk of 77.23%

</b>

```{r}
plot(MSFT.rr[2], MSFT.rr[1], pch=16, col="green", ylim=c(0, 1.30), xlim=c(0, .80), xlab = 'Annualized Risk', ylab = 'Annualized Return')
points(LMT.rr[2], LMT.rr[1], pch=16, col="blue")
points(ENPH.rr[2], ENPH.rr[1], pch=16, col="red")
```

</b>

## Part 1d:

</b>


```{r}
## SPY (S&P500 ETF)
SPY <- getSymbols("SPY",from=start, to=end, auto.assign=F)
SPY.close <- SPY[ ,4]
SPY.return <- dailyReturn(SPY.close)
```

</b>

```{r}
MSFT.cum.return <- cumprod(1 + MSFT.return)
LMT.cum.return <- cumprod(1 + LMT.return)
ENPH.cum.return <- cumprod(1 + ENPH.return)
SPY.cum.return <- cumprod(1 + SPY.return)
```

</b>

```{r}
plot(SPY.cum.return, type="l", ylim=c(0, 70), main = 'Cumulative Return in last 4 years')
lines(MSFT.cum.return, type="l", col="green")
lines(LMT.cum.return, type="l", col="blue")
lines(ENPH.cum.return, type="l", col="red")

```

</b>

```{r}
plot(ENPH.cum.return, type="l", col="red")
```
</b>

```{r}
plot(SPY.cum.return, type="l", ylim=c(0, 4), main = 'Cumulative Return in last 4 years')
lines(MSFT.cum.return, type="l", col="green")
lines(LMT.cum.return, type="l", col="blue")
```

</b>

# **Part2:**

</b>
</b>

## Part 2a:
</b>

```{r message=FALSE, warning=FALSE}
rf <- read.csv('F-F_Research_Data_Factors_daily.CSV', head=T, skip=3)
rf$date <- as.Date(rf$X, format='%Y%m%d')
attach(rf)
head(rf)
```

```{r}
rf <- na.omit(rf)
rf.avg <- mean(rf$RF)
rf.avg
```

  
```{r}
# Risk-free rate during the last 4 years
rf.new <- subset(rf, date >= start & date <= end)
rf.average <- mean(rf.new$RF)
```
  

Excess returns are the return earned by a stock (or portfolio of stocks) and the risk free rate, which is usually estimated using the most recent short-term government treasury bill.

For example, if a stock earns 15% in a year when the U.S. treasury bill earned 3%, the excess returns on the stock were 15% - 3% = 12%.

```{r}
SPY.excess <- 252 * (SPY.return - rf.avg)
MSFT.excess <- 252 * (MSFT.return - rf.avg)
LMT.excess <- 252 * (LMT.return - rf.avg)
ENPH.excess <- 252 * (ENPH.return - rf.avg)
head(SPY.excess)
```
```{r}
plot(SPY.excess$daily.returns, MSFT.excess$daily.returns, pch=20, type = 'p')
```

</b>

## Part 2b:

</b>

```{r}
MSFT.Alpha.Beta <- coef(lm(MSFT.excess ~ SPY.excess))
MSFT.Alpha.Beta
```


```{r}
LMT.Alpha.Beta <- coef(lm(LMT.excess ~ SPY.excess))
LMT.Alpha.Beta
```

```{r}
ENPH.Alpha.Beta <- coef(lm(ENPH.excess ~ SPY.excess))
ENPH.Alpha.Beta
```

```{r}
MSFT.Rsquare <- summary(lm(MSFT.excess ~ SPY.excess))$r.squared
MSFT.Rsquare
```
```{r}
LMT.Rsquare <- summary(lm(LMT.excess ~ SPY.excess))$r.squared
LMT.Rsquare
```

```{r}
ENPH.Rsquare <- summary(lm(ENPH.excess ~ SPY.excess))$r.squared
ENPH.Rsquare
```
</br>

|Underlying Stock | Ticker  | Alpha   | Beta    | Rsquare     |
|:----------------|:--------|:--------|:--------|:------------|
|Microsoft        |MFST     |0.75     |1.20     |0.68         |
|Lockheed Martin  |LMT      |-1.04    |0.64     |0.27         |
|Enphase Energy   |ENPH     |2.42     |1.58     |0.19         |